---
title: "Question 2"
output: html_document
---

To begin, we will import our dataset and all of the packages that will be used for this question. 
```{r}
library(tidyverse)

dat <- read.csv("./traits.csv")
summary(dat)
```
Next, we will select the columns that are going to be used to analyze our biological question. 
```{r}
dat <- dat %>% select(Coat_colour, Gender_status, Health, Agitated, NO_Agg, Train, AttenSeek)
summary(dat)
str(dat)
dat$Coat_colour <- as.factor(dat$Coat_colour)
dat$Gender_status <- as.factor(dat$Gender_status)
dat$Health <- as.factor(dat$Health)
```
As all of our non-predictor variables (coat colour, gender status, and health) are categorical and our predictor variables are numeric (agitated, non-owner directed agression, trainanbility, and attention seeking) we are going to create a loop that performs linear models and ANOVA tests for each combination of predictor and non predictor variables. 
```{r}
pred <- names(dat)[1:3]

#Agitated
resultsAgitated <- rep(NA, length(pred))
TukeyAgitated <- rep(NA, length(pred))

#Agression
resultsAgg <- rep(NA, length(pred))
TukeyAgg <- rep(NA, length(pred))

#Trainability
resultsTrain <- rep(NA, length(pred))
TukeyTrain <- rep(NA, length(pred))
#Attention Seeking
resultsAtt <- rep(NA, length(pred))
TukeyAtt <- rep(NA, length(pred))



for(i in 1:length(pred)){
  myformulaAgitated <- paste('Agitated~', pred[i])
  modAgitated <- aov(as.formula(myformulaAgitated), data = dat)
  resultsAgitated[i] <- summary(aov(modAgitated))
  TukeyAgitated[i] <- TukeyHSD(modAgitated)
  
  myformulaAgg <- paste('NO_Agg~', pred[i])
  modAgg <- aov(as.formula(myformulaAgg), data = dat)
  resultsAgg[i] <- summary(aov(modAgg))
  TukeyAgg[i] <- TukeyHSD(modAgg)

  
  myformulaTrain <- paste('Train~', pred[i])
  modTrain <- aov(as.formula(myformulaTrain), data = dat)
  resultsTrain[i] <- summary(aov(modTrain))
  TukeyTrain[i] <- TukeyHSD(modTrain)
  
  myformulaAtt <- paste('AttenSeek~', pred[i])
  modAtt <- aov(as.formula(myformulaAtt), data = dat)
  resultsAtt[i] <- summary(aov(modAtt))
  TukeyAtt[i] <- TukeyHSD(modAtt)
}

print(resultsAgitated)
print(TukeyAgitated)
print(resultsAgg)
print(TukeyAgg)
print(resultsTrain)
print(TukeyTrain)
print(resultsAtt)
print(TukeyAtt)



```

