---
title: "Question 2"
output: html_document
---

To begin, we will import our dataset and all of the packages that will be used for this question. 
```{r}
library(tidyverse)

dat <- read.csv("./traits.csv")
summary(dat)
```
Next, we will select the columns that are going to be used to analyze our biological question. 
```{r}
dat <- dat %>% select(Coat_colour, Gender_status, Health, Agitated, NO_Agg, Train, AttenSeek)
summary(dat)
```
As all of our non-predictor variables (coat colour, gender status, and health) are categorical and our predictor variables are numeric (agitated, non-owner directed agression, trainanbility, and attention seeking) we are going to create a loop that performs linear models and ANOVA tests for each combination of predictor and non predictor variables. 

Coat Colour
```{r}
pred <- names(dat)[4:7]

#Coat Colour
resultsCoat <- rep(NA, length(pred))
PvalCoat <- rep(NA, length(pred))

for(i in 1:length(pred)){
  myformula <- paste('Coat_colour~', pred[i])
  mod <- lm(as.formula(myformula), data = dat)
  resultsCoat[i] <- summary(aov(mod))
}
resultsCoat
```
Gender Status 
```{r}
resultsGender <- rep(NA, length(pred))

for(i in 1:length(pred)){
  myformula <- paste('Gender_status~', pred[i])
  mod <- lm(as.formula(myformula), data = dat)
  resultsGender[i] <- summary(aov(mod))
}
resultsGender
```
Health
```{r}
resultsHealth <- rep(NA, length(pred))
PvalHealth <- rep(NA, length(pred))

for(i in 1:length(pred)){
  myformula <- paste('Health~', pred[i])
  mod <- lm(as.formula(myformula), data = dat)
  resultsHealth[i] <- summary(aov(mod))
}
resultsHealth
```


```{r}
pred <- names(dat)[4:7]

#Coat Colour
resultsCoat <- rep(NA, length(pred))
PvalCoat <- rep(NA, length(pred))

#Gender
resultsGender <- rep(NA, length(pred))
PvalGender <- rep(NA, length(pred))

#Health Status
resultsHealth <- rep(NA, length(pred))
PvalHealth <- rep(NA, length(pred))


for(i in 1:length(pred)){
  myformulaCoat <- paste('Coat_colour~', pred[i])
  modCoat <- lm(as.formula(myformulaCoat), data = dat)
  resultsCoat[i] <- summary(aov(modCoat))
  
  myformulaHealth <- paste('Health~', pred[i])
  modHealth <- lm(as.formula(myformulaHealth), data = dat)
  resultsHealth[i] <- summary(aov(modHealth))
  
  myformulaGender<- paste('Gender_status~', pred[i])
  modGender <- lm(as.formula(myformulaGender), data = dat)
  resultsGender[i] <- summary(aov(modGender))
}

print(resultsCoat)
print(resultsHealth)
print(resultsGender)
```

#vvvv Rough Tukey attempt vvvv

```{r}
lmCoat<-lm(Coat_colour ~ as.factor(Agitated) + as.factor(NO_Agg) + as.factor(Train) + as.factor(AttenSeek), data = dat)
aovCoat<-aov(lmCoat)
```

```{r}
summary(aovCoat)
```

```{r}
(TukeyHSD(aovCoat))
```

