---
title: "Question 2"
author: "Jordan Rutherford"
date: "12/03/2021"
output: html_document
---

### Introduction

This script will evaluate the influence of age on the expression of different personality traits in Labrador Retrievers. Age is measured in days.

```{r}
traits_dat <- read.csv("traits.csv") #loading in my data
traits_dat$Age <- as.numeric(traits_dat$Age)
```

```{r}
#library load
library(ggplot2)
library(cowplot)
```


### Creating graphs to evaluate the distribution of age across each behavioural trait:

```{r fig.height = 7, warning = FALSE}
#this loop creates a scatter plot for each behavioural trait vs age
for (i in 1:length(traits_dat)) {
  plot <- ggplot(data = traits_dat,
                 aes_string(x = "Age", y = names(traits_dat[i]))) + #plot age on x and behavioural traits on y
    geom_point(alpha = 0.1) + #reduce opacity to visualize areas of high density
    theme_classic()
  
  #assign names of plots based on trait names to an object
  assign(x = paste0(names(traits_dat)[i], "_plot"), plot)
}

#view these generated plots in a grid
plot_grid(Agitated_plot, AttenSeek_plot, Bark_plot, Excite_plot, Fetch_plot, HO_Fear_plot, NO_Agg_plot, NoiseF_plot, O_Agg_plot, SepAnx_plot, Train_plot, UnBeh_plot, ncol = 3)
```


### Analyzing the data

We will be determining whether age has a significant effect on the expression of each of the behavioural traits in question using a one-way ANOVA for each group. 

```{r, warning = FALSE}
#create subset of the behaviours we are analyzing
behaviours <- names(traits_dat)[8:19]

#create a blank data frame to hold the data for our analyses
age_data <- data.frame(matrix(nrow = 12, ncol = 4))
names(age_data) <- c("Behavioural_Trait", "F_value", "Adj_R2", "P_value")

#loop to iterate through each trait and evaluate the influence of age
for (i in 1:length(behaviours)) {
  #create linear models to plug into analyses
  lms <- paste(behaviours[i], "~ Age")
  #perform linear analyses using lms object
  analyses <- summary(lm(as.formula(lms), data = traits_dat))
  
  #store behaviour names
  age_data$Behavioural_Trait[i] <- behaviours[i]
  #store F values
  age_data$F_value[i] <- analyses$fstatistic
  #store adjusted R2 values
  age_data$Adj_R2[i] <- analyses$adj.r.squared
  #compute and store p values
  age_data$P_value[i] <- pf(analyses$fstatistic[1],
                            analyses$fstatistic[2],
                            analyses$fstatistic[3],
                            lower.tail = F)
  
}

#print linear regression results
print(age_data)
```


### Graphing the linear regression results

```{r}
#graph R2 values for influence of age vs behavioural traits
ggplot(age_data, aes(x = Behavioural_Trait, y = Adj_R2)) +
  geom_bar(stat = "identity") +
  theme_classic()
```


```{r}
#graph p values for influence of age vs behavioural traits
ggplot(age_data, aes(x = Behavioural_Trait, y = P_value)) +
  geom_bar(stat = "identity") +
  theme_classic() +
  geom_hline(yintercept = 0.05, colour = "red") #add line at p = 0.05 to show significant values
```















