---
title: "Question 2"
output: html_document
---

To begin, we will import our dataset and all of the packages that will be used for this question. 
```{r}
library(tidyverse) #loading tidyverse 
library(cowplot) #loading cowplot
 
dat <- read.csv("./traits.csv") #loading our data 
summary(dat)
```
Next, we will select the columns that are going to be used to analyze our biological question. 
```{r}
dat <- dat %>% select(Coat_colour, Gender_status, Health, Agitated, NO_Agg, Train, AttenSeek)
summary(dat) #selecting our data columns 
str(dat) #checking the structure of our data 
dat$Coat_colour <- as.factor(dat$Coat_colour) #changing coat colour to a factor
dat$Gender_status <- as.factor(dat$Gender_status) #changing gender status to a factor
dat$Health <- as.factor(dat$Health) #changing health status to a factor 
```
First, we will visualize the data utilizing boxplots. Let's take a look at the effect of coat colour on predictor variables. 
```{r fig.height=10, fig.width=12}
#adding a theme to make our plots look publication worthy
mytheme <- theme(
  axis.ticks = element_line(size=1, colour="black"),
  axis.text.x = element_text(size=16, colour="black"),
  axis.text.y = element_text(size=16,, colour="black"),
  axis.title.x = element_text(size=20, colour="black", vjust=0.5),
  axis.title.y = element_text(size=20, colour="black", vjust=+2),
  panel.background = element_rect(fill="white", colour="black"),
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
)

#Coat Colour plots against our predictor variables 
CAgg <- ggplot(data = dat) +
  geom_boxplot(aes(x = Coat_colour, y = Agitated, fill = Coat_colour)) +
  xlab("Coat Colour") +
  ylab("Aggitated") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(1,5), breaks = seq(1,5,1)) +
  scale_x_discrete(labels = c("Black","Yellow","Chocolate")) +
  mytheme

CNO <- ggplot(data = dat) +
  geom_boxplot(aes(x = Coat_colour, y = NO_Agg, fill = Coat_colour)) +
  xlab("Coat Colour") + 
  ylab("Non-owner Directed Agression") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(1,5), breaks = seq(1,5,1)) +
  scale_x_discrete(labels = c("Black","Yellow","Chocolate")) +
  mytheme

CT <- ggplot(data = dat) +
  geom_boxplot(aes(x = Coat_colour, y = Train, fill = Coat_colour)) +
  xlab("Coat Colour") +
  ylab("Trainability") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(1,5), breaks = seq(1,5,1)) +
  scale_x_discrete(labels = c("Black","Yellow","Chocolate")) +
  mytheme

CAtt <- ggplot(data = dat) +
  geom_boxplot(aes(x = Coat_colour, y = AttenSeek, fill = Coat_colour)) +
  xlab("Coat Colour") +
  ylab("Attention Seeking") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(1,5), breaks = seq(1,5,1)) +
  scale_x_discrete(labels = c("Black","Yellow","Chocolate")) +
  mytheme

pdf(file = "./Question_2_Figures/Coat_Colour.pdf", height = 10, width = 10)
(CColour <- plot_grid(CAgg, CNO, CT, CAtt, labels = "AUTO")) #creating a 4 plot graph
dev.off()
```
Next, lets take a look at the effect of gender status on predictor variables. 
```{r fig.height=10, fig.width=20}
#Gender status plots against our predictor variables 
GAgg <- ggplot(data = dat) +
  geom_boxplot(aes(x = Gender_status, y = Agitated, fill = Gender_status)) +
  xlab("Gender Status") +
  ylab("Aggitated") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(1,5), breaks = seq(1,5,1)) +
  scale_x_discrete(labels = c("Male Entire","Male Neutered",
          "Female Entire", "Female Neutered", "Missing")) +
  mytheme

GNO <- ggplot(data = dat) +
  geom_boxplot(aes(x = Gender_status, y = NO_Agg, fill = Gender_status)) +
  xlab("Gender Status") + 
  ylab("Non-owner Directed Agression") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(1,5), breaks = seq(1,5,1)) +
  scale_x_discrete(labels = c("Male Entire","Male Neutered",
          "Female Entire", "Female Neutered", "Missing")) +
  mytheme

GT <- ggplot(data = dat) +
  geom_boxplot(aes(x = Gender_status, y = Train, fill = Gender_status)) +
  xlab("Gender Status") +
  ylab("Trainability") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(1,5), breaks = seq(1,5,1)) +
  scale_x_discrete(labels = c("Male Entire","Male Neutered",
          "Female Entire", "Female Neutered", "Missing")) +
  mytheme

GAtt <- ggplot(data = dat) +
  geom_boxplot(aes(x = Gender_status, y = AttenSeek, fill = Gender_status)) +
  xlab("Gender Status") +
  ylab("Attention Seeking") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(1,5), breaks = seq(1,5,1)) +
  scale_x_discrete(labels = c("Male Entire","Male Neutered",
          "Female Entire", "Female Neutered", "Missing")) +
  mytheme

pdf(file = "./Question_2_Figures/Gender_Status.pdf", height = 10, width = 20)
(GStatus <- plot_grid(GAgg, GNO, GT, GAtt, labels = "AUTO")) #creating a 4 plot graph
dev.off()
```
Finally, we will look at the effect of health on predictor variables
```{r fig.height=10, fig.width=12}
#Health Status plots against our predictor variables 
HAgg <- ggplot(data = dat) +
  geom_boxplot(aes(x = Health, y = Agitated, fill = Health)) +
  xlab("Health") +
  ylab("Aggitated") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(1,5), breaks = seq(1,5,1)) +
  scale_x_discrete(labels = c("Health Problems", "No Health Problems")) +
  mytheme

HNO <- ggplot(data = dat) +
  geom_boxplot(aes(x = Health, y = NO_Agg, fill = Health)) +
  xlab("Health") + 
  ylab("Non-owner Directed Agression") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(1,5), breaks = seq(1,5,1)) +
  scale_x_discrete(labels = c("Health Problems", "No Health Problems")) +
  mytheme

HT <- ggplot(data = dat) +
  geom_boxplot(aes(x = Health, y = Train, fill = Health)) +
  xlab("Health") +
  ylab("Trainability") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(1,5), breaks = seq(1,5,1)) +
  scale_x_discrete(labels = c("Health Problems", "No Health Problems")) +
  mytheme

HAtt <- ggplot(data = dat) +
  geom_boxplot(aes(x = Health, y = AttenSeek, fill = Health)) +
  xlab("Health") +
  ylab("Attention Seeking") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(1,5), breaks = seq(1,5,1)) +
  scale_x_discrete(labels = c("Health Problems", "No Health Problems")) +
  mytheme

pdf(file = "./Question_2_Figures/Health_Status.pdf", height = 10, width = 10)
(HStatus <- plot_grid(HAgg, HNO, HT, HAtt, labels = "AUTO")) #creating a 4 plot graph 
dev.off()
```
As all of our non-predictor variables (coat colour, gender status, and health) are categorical and our predictor variables are numeric (agitated, non-owner directed agression, trainanbility, and attention seeking) we are going to create a loop that performs ANOVA tests for each combination of predictor and non predictor variables. 
```{r}
npred <- names(dat)[1:3] #creating a vector of column names 

#Agitated vectors for results and TukeyHSD
resultsAgitated <- rep(NA, length(npred))
TukeyAgitated <- rep(NA, length(npred))

#Agression vectors for results and TukeyHSD
resultsAgg <- rep(NA, length(npred))
TukeyAgg <- rep(NA, length(npred))

#Trainability vectors for results and TukeyHSD
resultsTrain <- rep(NA, length(npred))
TukeyTrain <- rep(NA, length(npred))

#Attention Seeking vectors for results and TukeyHSD
resultsAtt <- rep(NA, length(npred))
TukeyAtt <- rep(NA, length(npred))



for(i in 1:length(npred)){
  myformulaAgitated <- paste('Agitated~', npred[i]) #creating a formula for our anova test 
  modAgitated <- aov(as.formula(myformulaAgitated), data = dat) #applying our anova test 
  resultsAgitated[i] <- summary(aov(modAgitated)) #outputting our results 
  TukeyAgitated[i] <- TukeyHSD(modAgitated) #outputting TukeyHSD 
  
  myformulaAgg <- paste('NO_Agg~', npred[i]) #creating a formula for our anova test
  modAgg <- aov(as.formula(myformulaAgg), data = dat) #applying our anova test
  resultsAgg[i] <- summary(aov(modAgg)) #outputting our results 
  TukeyAgg[i] <- TukeyHSD(modAgg) #outputting TukeyHSD

  
  myformulaTrain <- paste('Train~', npred[i]) #creating a formula for our anova test
  modTrain <- aov(as.formula(myformulaTrain), data = dat) #applying our anova test 
  resultsTrain[i] <- summary(aov(modTrain)) #outputting results 
  TukeyTrain[i] <- TukeyHSD(modTrain) #outputting TukeyHSD
  
  myformulaAtt <- paste('AttenSeek~', npred[i]) #creating a formula for our anova test 
  modAtt <- aov(as.formula(myformulaAtt), data = dat) #applying our anova test 
  resultsAtt[i] <- summary(aov(modAtt)) #outputting results 
  TukeyAtt[i] <- TukeyHSD(modAtt) #outputting TukeyHSD
}

print(resultsAgitated)
print(TukeyAgitated)
print(resultsAgg)
print(TukeyAgg)
print(resultsTrain)
print(TukeyTrain)
print(resultsAtt)
print(TukeyAtt)
```


