---
title: "Question 2"
output: html_document
---

To begin, we will import our dataset and all of the packages that will be used for this question. 
```{r}
library(tidyverse)

dat <- read.csv("./traits.csv")
summary(dat)
```
Next, we will select the columns that are going to be used to analyze our biological question. 
```{r}
dat <- dat %>% select(Coat_colour, Gender_status, Health, Agitated, NO_Agg, Train, AttenSeek)
summary(dat)
```
As all of our non-predictor variables (coat colour, gender status, and health) are categorical and our predictor variables are numeric (agitated, non-owner directed agression, trainanbility, and attention seeking) we are going to create loops that performs linear models and ANOVA tests for each combination of predictor and non predictor variables. 

Coat Colour
```{r}
pred <- names(dat)[4:7]

#Coat Colour
resultsCoat <- rep(NA, length(pred))
PvalCoat <- rep(NA, length(pred))

for(i in 1:length(pred)){
  myformula <- paste('Coat_colour~', pred[i])
  mod <- lm(as.formula(myformula), data = dat)
  resultsCoat[i] <- summary(aov(mod))
}
resultsCoat
```
Gender Status 
```{r}
resultsGender <- rep(NA, length(pred))

for(i in 1:length(pred)){
  myformula <- paste('Gender_status~', pred[i])
  mod <- lm(as.formula(myformula), data = dat)
  resultsGender[i] <- summary(aov(mod))
}
resultsGender
```
Health
```{r}
resultsHealth <- rep(NA, length(pred))
PvalHealth <- rep(NA, length(pred))
for(i in 1:length(pred)){
  myformula <- paste('Health~', pred[i])
  mod <- lm(as.formula(myformula), data = dat)
  resultsHealth[i] <- summary(aov(mod))
}
resultsHealth
```

