---
title: "sex_neuter_analysis"
author: "Alex Tsiofas"
date: "11/03/2021"
output: html_document
---
## Intro
This script will visualize and determine if any of the
Neutering categories - Male entire (n = 451), male neutered
(n = 59), female entire (n = 1028), female neutered (n = 426)- 
are significantly different from the others in any of the 
behavioural traits

## Loading Libraries and Importing the Traits csv file from Git
```{r warning = F, message=F}
library(ggplot2)
library(dplyr)
library(cowplot)

traits_dat <- read.csv("traits.csv")

traits_dat$Gender_status <- as.factor(traits_dat$Gender_status)
```

## Plotting the Data
```{r fig.height= 8, fig.width = 8, warning=F}
#loop that creates a boxplot for each of the traits in the list, 
#plotted against gender/neutering status
for (i in 8:19){
  #plot with gender status on x and each trait on y
  plot1 <- ggplot(data = subset(traits_dat, !is.na(Gender_status)), 
                  aes_string(x = "Gender_status", y = names(traits_dat[i])))+
    geom_boxplot()+
    scale_x_discrete(labels = c("Male Entire", "Male Neutered", 
                                "Female Entire", "Female Neutered"))+
    theme_classic()
  #assigning the plot to an object with the behavioural trait name
  assign(x = paste0(names(traits_dat)[i], "_plot"), plot1)
}

plot_grid(Agitated_plot, AttenSeek_plot, Bark_plot, Excite_plot, Fetch_plot, HO_Fear_plot, NO_Agg_plot,
          NoiseF_plot, O_Agg_plot, SepAnx_plot, Train_plot, UnBeh_plot,
          ncol = 3)
```

## Analysis
Within each behaviour we are comparing whether the mean ranking of the behaviour significantly differs between  the gender/neuter status groups. We use an ANOVA to compare these group means
```{r warning = F}
#making a list of the behaviour column names
behaviours1 <- names(traits_dat)[8:19]

#initiating a data frame to hold the anova statistics
sexneuter_anovas <- data.frame(matrix(nrow = 12, ncol = 3))
names(sexneuter_anovas) <- c("Behaviour", "F_Statistic", "P_Value")

for (i in 1:length(behaviours1)){
  #looking at how gender status impacts each behaviour
  rel1 <- paste(behaviours1[i], '~ Gender_status')
  #running the anova with the specified function
  analysis1 <- aov(as.formula(rel1), data = traits_dat)
  #assigning the analysis results to an appropriate variable name
  sexneuter_anovas$Behaviour[i] <- behaviours1[i]
  sexneuter_anovas$F_Statistic[i] <- summary(analysis1)[[1]][["F value"]]
  sexneuter_anovas$P_Value[i] <- summary(analysis1)[[1]][["Pr(>F)"]]
}

print(sexneuter_anovas)
```

